---
title: "prelim analyses"
author: "Alejandra Garcia Isaza"
date: "2023-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal(base_size = 15))
```

# Loading the dataset
```{r}
d <- read_sav(here("nopublish", "LEA Parent W1.2 FINAL.sav")) %>%
  arrange(P0001.1) 


d1 <- d %>% 
  select(-2:-11) %>% 
  janitor::clean_names() %>%
  rename(c("id" = "p0001_1"),
         c("condition" = "e0002_1"),
         c("site" = "e0003_1")) %>%
  mutate(id = case_when(id == "XXXX" ~ "PE27", # Assigning id PE27 to row id under XXXX - confirmed with response_id from qualtrics & logs
                        TRUE ~ as.character(id))) %>%
  mutate(condition = case_when(id == "PE27" ~ "0", # Assigning control condition to id PE27 - participant was not part of intervention groups
                        TRUE ~ as.character(condition))) %>%
  mutate(site = case_when(id == "PE27" ~ "1", # Assigning site to Oregon to id PE27 
                        TRUE ~ as.character(site))) %>%
  arrange(id)

# d1 <- d1 %>%
#   mutate(condition = factor(condition),
#          condition = fct_recode(condition,
#                            "Control" = "0",
#                            "Intervention" = "1")) %>%
#   mutate(site = factor(site),
#          site = fct_recode(site,
#                            "Oregon" = "1",
#                            "Florida" = "2"))
```

```{r}
view_df(d)
view_df(d1)
```
To compare groups on demo vars, I need to separate mothers from fathers, but they don't have the same family id. Here is what I was able to identify:

OREGON
couples that I remember:
EUGENE (INT)
Mother: PE05, Father: PE16
Mother: PE13, Father: PE08

NOT SURE, SAME LAST NAME, AND DAUGHTER NAME:
EUGENE (CTRL)
Mother: PE20, Father: PE22 

FLORIDA
couples that I remember:
Miami (INT)
Mother: PM14, Father: PM17

SAME ADDRESS, SAME DAUGHTER NAME:
Miami (CTRL)
Mother: PM16, Father: PM15

approach: include all parents, acknowledge in paper and ask Seth

----
location_rc_1: location recode

p8011_1	What is the highest level of education you
completed? (Mark only one)

p8013_1	household income

```{r}
# ggplot(d1, aes(location_rc_1)) +
#   geom_bar() 
#   # facet_wrap(~ condition)
  
ggplot(d1, aes(location_rc_1)) +
  geom_histogram()
```
```{r}
d1 %>%
  group_by(location_rc_1) %>%
  summarise(avg_income = mean(p8013_1))
```

```{r}
# total = 39 participants
d_plot_eug <- d1 %>%
  filter(location_rc_1 == 1) # 22 participants

d_plot_flo <- d1 %>%
  filter(location_rc_1 == 2) # 10 participants

d_plot_other <- d1 %>%
  filter(location_rc_1 == 3) # 5 participants

# 2 participants with NA in variable are from Orlando
```
is it best to group by west coast/east coast?

Approach: report that roughly participants come from west coast and east coast with a little bit of geographic variability. The purpose was to identify variability by region, but there was the pandemic and people moved or people who were referred gave info for people who met criteria but moved to other cities in the general west coas/east coast

move PE19 to west coast group

```{r}
ggplot(d_plot_eug, aes(p8011_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  labs(title = "Parent's Ed. level Oregon",
       y = "Count",
       x = "Ed. level")

ggplot(d_plot_eug, aes(p8013_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)) +
  labs(title = "Parent's Income level Oregon",
       y = "Count",
       x = "Income level")
```

```{r}
ggplot(d_plot_flo, aes(p8011_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  labs(title = "Parent's Ed. level Florida",
       y = "Count",
       x = "Ed. level")

ggplot(d_plot_flo, aes(p8013_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)) +
  labs(title = "Parent's Income. level Florida",
       y = "Count",
       x = "Income level")
```

```{r}
ggplot(d_plot_other, aes(p8011_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  labs(title = "Parent's Ed. level Other",
       y = "Count",
       x = "Ed. level")

ggplot(d_plot_other, aes(p8013_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)) +
  labs(title = "Parent's Income level Other",
       y = "Count",
       x = "Income level")
```


### Plots by Region ###


```{r}
# total = 39 participants
d_plot_west <- d1 %>%
  filter(site == 1) # 23 participants (12 int & 11 ctrl)

d_plot_east <- d1 %>%
  filter(site == 2)  # 16 participants (7 int & 9 ctrl)
```

```{r}
ggplot(d_plot_west, aes(p8011_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  labs(title = "Parent's Ed. level West Coast",
       y = "Count",
       x = "Ed. level")

ggplot(d_plot_west, aes(p8013_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)) +
  labs(title = "Parent's Income level West Coast",
       y = "Count",
       x = "Income level")
```

```{r}
ggplot(d_plot_east, aes(p8011_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) +
  labs(title = "Parent's Ed. level East Coast",
       y = "Count",
       x = "Ed. level")

ggplot(d_plot_east, aes(p8013_1)) +
  geom_bar() +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)) +
  labs(title = "Parent's Income level East Coast",
       y = "Count",
       x = "Income level")
```
```{r}
ggplot(d_plot_west, aes(p5001_2)) +
  geom_bar() +
  labs(title = "Any Alcoholic drink - West Coast")

ggplot(d_plot_east, aes(p5001_2)) +
  geom_bar() +
  labs(title = "Any Alcoholic drink - East Coast")
```


p5004_2

For parents: check if there are differences between sites on key variables

IV: Site (West coast/east coast) - grouping var.
DV: income level range - grouping var.
Analysis: Chi-square

IV: Site (West coast/east coast) - grouping var.
DV: educational level - grouping var.
Analysis: Chi-square

IV: Site (West coast/east coast) - grouping var.
DV: acculturation - grouping var.
Analysis:


FOR EAs

gender, binge drinking, acculturation, language, attending college, # of jobs, differences, between EAs by site. 
